Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-08 paper/paper.tex

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
%\usepackage{algorithmic}

@@

\newcommand{\newC}{C$^-$\xspace}

@@

\begin{figure}
\centering
\begin{minipage}{0.9\textwidth}
 \begin{lstlisting}[mathescape,language=C,morekeywords={function}]
function synth(inputs) {
  $(i_1, \ldots, i_N)$ = inputs
  query := $\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$
  result := decide(query)

  if (result.satisfiable) {
    return result.model
  } else {
    return unsatisfiable
  }
}

function verif(P) {
  query := $\exists x . \lnot \sigma(x, P(x))$
  result := decide(query)

  if (result.satisfiable) {
    return result.model
  } else {
    return valid
  }
}

function refinement_loop() {
  inputs := $\emptyset$

  while (true) {
    candidate := synth(inputs)

    if (candidate = unsatisfiable) {
      return unsatisfiable
    }

    res := verif(candidate)

    if (res = valid) {
      return candidate
    } else {
      inputs := inputs $\cup$ res
    }
  }
 \end{lstlisting}
 \end{minipage}

 \caption{Abstract refinement algorithm
 \label{fig:abstract-refinement-code}}
\end{figure}

@@

Each opcode requires $\log_2 I$ bits to encode, where $I$ is the number
of instruction types in $\mathcal{L}$.  Each operand can be encoded using
$\log_2 w$ bits, where $w$ is the $\mathcal{L}$-machine word, plus one
bit to specify whether the operand is a register name or an immediate constant.
One instruction can therefore be encoded using $\log_2 I + 2w + 2$ bits.
For an $n$ instruction program, we need $$n \log_2 I + 2nw + 2n$$ bits to encode
the entire program.

If we instead limit the number of constants that can appear in the program,
our operands can be encoded using fewer bits.  For an $n$ instruction program
using $k$ constants and taking $a$ arguments as inputs, each operand can refer
to a program argument, the result of a previous instruction or a constant.
This can be encoded using $\log_2 (k+a+n-1)$ bits, which means each instruction
can be encoded in $\log_2 I + \log_2 (k + a + n - 1)$ and the full program
needs $$n \log_2 I + n \log_2 (k + a + n - 1) + kw$$ bits to encode.

>>>>

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}

@@

\newcommand{\newC}{C$^-$\xspace}
\newcommand*\Let[2]{\State #1 $\gets$ #2}

@@

\begin{algorithm}
 \caption{Abstract refinement algorithm
 \label{fig:abstract-refinement-code}}

 \begin{algorithmic}[1]
\Statex
\Function{synth}{inputs}
  \Let{$(i_1, \ldots, i_N)$}{inputs}
  \Let{query}{$\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{unsatisfiable}
  \EndIf
\EndFunction
\Statex
\Function{verif}{P}
  \Let{query}{$\exists x . \lnot \sigma(x, P(x))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{valid}
  \EndIf
\EndFunction
\Statex
\Function{refinement loop}{}
  \Let{inputs}{$\emptyset$}
  \Loop
    \Let{candidate}{synth(inputs)}
    \If{candidate = unsatisfiable}
      \State \Return{unsatisfiable}
    \EndIf
    \Let{res}{verif(candidate)}
    \If{res = valid}
      \State \Return{candidate}
    \Else
      \Let{inputs}{inputs $\cup$ res}
    \EndIf
  \EndLoop
\EndFunction
 \end{algorithmic}
\end{algorithm}

@@

Each opcode requires $\lceil \log_2 I \rceil$ bits to encode, where $I$ is the number
of instruction types in $\mathcal{L}$.  Each operand can be encoded using
$\log_2 w$ bits, where $w$ is the $\mathcal{L}$-machine word, plus one
bit to specify whether the operand is a register name or an immediate constant.
One instruction can therefore be encoded using $\lceil \log_2 I \rceil + 2w + 2$ bits.
For an $n$ instruction program, we need $$\lceil n \log_2 I \rceil + 2nw + 2n$$ bits to encode
the entire program.

If we instead limit the number of constants that can appear in the program,
our operands can be encoded using fewer bits.  For an $n$ instruction program
using $k$ constants and taking $a$ arguments as inputs, each operand can refer
to a program argument, the result of a previous instruction or a constant.
This can be encoded using $\lceil \log_2 (k+a+n-1) \rceil$ bits, which means each instruction
can be encoded in $\lceil \log_2 I \rceil + \lceil \log_2 (k + a + n - 1) \rceil$ and the full program
needs $$\lceil n \log_2 I \rceil + \lceil n \log_2 (k + a + n - 1) \rceil + kw$$ bits to encode.
